DRAFT\\Satisfiability Modulo Theories Competition (SMT-COMP) 2014: Rules and Procedures

### 2014-05-07 rules/rules14.tex
Next draft of the 2014 rules.

\title{Satisfiability Modulo Theories Competition (SMT-COMP) 2014: Rules and 
Procedures}

@@

\author{
David R. Cok \\
GrammaTech, Inc. \\
Ithaca, NY (USA) \\
cok@frontiernet.net \\
\and
David Deharbe \\
Federal University of Rio Grande do Norte \\
Brazil \\
David.Deharbe@loria.fr \\
\and
Tjark Weber \\
Uppsala University, \\
Sweden, \\
tjark.weber@it.uu.se \\ 
}

@@

\noindent Comments on this document should be emailed to the smtcomp mailing
list or, if necessary, directly to the organizers.

\section{Important dates and information}
\begin{itemize}
\item May 15 - deadline for new benchmarks
\item May 25 - deadline for solver entrants
\item June 1 - deadline for final versions of solvers (but no new entrants after May 25).
\item June 17 - SMT workshop; completion of competition.
\item SMT COMP mailing list: smt-comp@cs.nyu.edu
\item signup site for mailing list: \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\item competition web site: \url{www.smtcomp.org}  or \url{smtcomp.sourceforge.net/2014}
\end{itemize}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR,TBD-2012,TBD-2013}).
SMT-COMP 2014 is affiliated with the SMT workshop (\url {http://smt2014.it.uu.se/}) at the co-located meetings of CAV, IJCAR, and SAT at FLoC 2014 (\url{http://vsl2014.at/}); the competition is part of the
Olympic Games (\url{http://vsl2014.at/olympics/}).

@@

SMTCOMP 2014 will have two tracks: the conventional main track and a repeat of 2012's application track.
Within a track there are one or more divisions, where each division
uses benchmarks from a specific SMT-LIB logic (or group of logics).

The rest of this document, revised from the previous year's
version\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.},
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from last competition are the following:
\begin{itemize}
\item The competition will be run with StarExec. Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [ The demonstration parallel solvers, unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.]
\item In the main track, we will measure comparative performance within all divisions to which sufficient solvers are submitted.

@@

\item In addition to the traditional SMTCOMP winners, we will recognize winners in other categories, as described below. 
Also, importantly, the Olympic Games has an award ceremony (with actual medals). A selection of the SMTCOMP winners (as described below) will be awarded Olympic Games medals.

@@

All results of the competition will be made public. However, 
winners will be declared only for divisions in which at least two solvers are submitted.
Although the organizers may add other solvers for comparison purposes, only solvers that
are explicitly submitted by their authors' are eligible to be designated as winners.
If necessary, and based on final submissions, the organizers may elect to
combine competition divisions to make them competitive (such decisions will
be made with input from the community).

@@

For participation in SMT-COMP, a solver must be
uploaded to Star-Exec, made publicly available, and the organizers informed of its presence {|em and thte logic divisions in which it is participating}.

A submission must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

\header{System description.} %
As part of a submission,
SMT-COMP entrants are encouraged to provide a short (1--2 pages) description of
the system.  This should include a list of all authors of the system
and their present institutional affiliations.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool, but this is
optional.

\header{Other systems.} %
As in previous years, due to limitations on
computational resources, the organizers reserve the right not to
accept multiple versions of the same solver (defined as sharing 50\%
or more of its source code).  The organizers reserve the right to
include other systems of interest (such as entrants in previous competitions), to the
competition.

\header{Wrapper tools.} %
A \emph{wrapper tool} is defined as any tool
which calls one or more SMT solvers not written by the author of the wrapper
tool.  The other solvers are is called the \emph{wrapped} tools.  
A wrapper tool must explicitly acknowledge any tools that it wraps.
Its system description should make clear the technical innovations  by which the
wrapper tool expects to improve on the wrapped tools.

@@

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via StarExec and email to the organizers by 7pm, Eastern U.S.
time, May~25, 2014.  At that time new entrants will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until 7pm, Eastern U.S. time, June~1, 2014.  We strongly encourage
participants to use this grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using StarExec after the original deadline on May~25.

@@

\paragraph{Parallel track}
StarExec reports both cpu time and wall clock time. The main competition will be measured by cpu time taken across all processes and threads that are spawned by the starting process. For single-process executions the wall-clock time is typically slightly larger than the cpu time. Multi-threaded solvers will see reduced wall-clock times. While we will not run an explicit parallel track in 2014. We will recognize the solvers with the best wall-clock time in each division.

@@

Solvers will be publicly evaluated in the following tracks. In exhibition divisions of these tracks, solvers are evaluated and results posted, but no awards are announced. The other divisions are competition divisions; if there are sufficient entrants, a winner will be announced.
\begin{itemize}
\item a main track: sequential execution evaluated separately on benchmarks from each of the logics for which there are benchmarks in StarExec.
\item application track: evaluation on command scripts (all divisions with sufficient benchmarks and entrants are competitive)
\end{itemize}

@@

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the weeks leading up to SMT 2014, from about June~2 to the July~17.  Results will be
announced at SMT 2014 and in an awards ceremony scheduled by the Olympic Games coordinator.
  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

@@

\header{Input and Output.}
%
Participating solvers must read a single benchmark script (defined
below, not part of the 2.0 standard), presented on its standard input
channel. The script is in the concrete syntax of the SMT-LIB format,
version 2.0.   A
\textbf{benchmark script} is just a text file containing a sequence of SMTLIBv2 commands in which:

@@

\noindent The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated during their submission to SMT-Exec.
Example benchmark scripts for several Problem Divisions are reported in the Appendix.
Note that they are provided for illustrative purposes only: 
please refer to the SMT-LIB format specification 
and the above definition of benchmark script for the official specification 
of the input format for SMT-COMP.

@@

The main track competition uses a StarExec postprocessor (SMT2 results conserv) to accumulate the results.

Solvers that register in non-competition logic divisions (cf. section \ref{sec:theories})
will be run in an exhibition mode: results will be generated and publicly
posted, but no awards will be announced.

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given formula, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of wall clock time and 6000 seconds of CPU time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
% TBD- TODO - how to constrain solvers to single processes?
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.  A
timeout scores the same as if the output is ``\texttt{unknown}''.

@@

\medskip
Solvers must respond immediately to the commands sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands, and with the defined responses for \akey{get-unsat-core} and \akey{get-proof} commands.

\header{Timeouts.}
A time limit is set for the whole execution of each application
benchmark (consisting of multiple queries).  We anticipate the
timeout to be around 25~minutes (as it has been in the past).\footnote{The timeout may be adjusted once we have experience with the abilities and resources of the StarExec service.}

@@

In 2012 the competition ran only some logics as competitive divisions.
In 2014 we will revert to previous practice and run all logics as competitive divisions, if there are multiple solvers submitted.

The organizers reserve the right to include in the benchmark population for a given logic benchmarks
from subsumed logics. For example, QF\_UFLIA may include some samples from QF\_LIA, and QF\_UFLRA may include some from QF\_LRA.
We will not use benchmarks from QF\_UF in other divisions.

\subsection{Main and parallel tracks}

\header{Benchmark sources.} %
Benchmark formulas for these divisions
will be drawn from the SMT-LIB library. The deadline for new benchmarks is
May 15. The organizers will be checking and curating these until June 1;
the organizers reserve the option to exclude new benchmarks if any prove problematic for some reason.
SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.} % 
A first release of the competition
benchmarks will be made available by May 15, 2014.
A second and almost final release will be available on June 1st. 
No additional
benchmarks will be added after this date, but benchmarks may be
modified or removed to fix possible bugs or other issues, or to adjust their difficulty score (see below). 
The final release that will be used for the competition will be posted on June
1. The set of selected benchmarks will be published when the
competition begins.

@@

Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (\ie\ benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (\ie\ benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $2^{31}-1$).

@@

The benchmarks are assigned a difficulty. For 2014, the difficulty measure is the
cpu time taken by the best performing solver in the 2013 SMTEVAL. These times will be posted 
as soon after May 15 as possible.

@@

\item \textbf{Retire very easy benchmarks.} %
  Eliminate benchmarks for which all of the current solvers in SMTEVAL-2103
  solved the benchmark in under 5 seconds,
  \emph{unless} doing so reduces the pool of
    benchmarks for the division to less than~300.

@@

Then, for each division, the benchmarks are divided into quintiles according to the difficulty measure.
Then a random selection of N competition benchmarks is made as follows:
\begin{itemize}
\item N*40\% are taken from the top quintile, if sufficient benchmarks are available
\item additional benchmarks are taken from the second quintile, to produce N*55\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the third quintile, to produce N*70\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the fourth quintile, to produce N*85\% benchmarks, if sufficient benchmarks are available
\item any remaining needed benchmarks are taken from the bottom quintile.
\end{itemize}
Within each quintile selection, if sufficient benchmarks are available, 85\% will be chosen from the industrial benchmarks,

Since the organizers at this point are unsure how long a set of benchmarks may take (which will depend also on the number of solvers submitted), the competition will be run in {\em heats}. Some number $N$ of benchmarks will be selected and the competition run. If there is time, an additional $N_1$ benchmarks will be run, etc., until the end of the competition is reached.

%Before the selection process, each benchmark will be assigned a
%\emph{difficulty:} a number between 0.0 and 5.0 inclusive, calculated as in previous years.  The
%difficulty for a particular benchmark will be assigned by running SMT
%solvers from previous competitions that finished in good
%standing and using the formula:
%
%\[\mathrm{difficulty} = {5\cdot\ln\!\left(1+A^2\right) \over \ln\!\left(1+30^2\right)}\]
%
%\noindent
%where $A$ is the average time for the solvers to correctly solve the
%instance (in minutes).  This computation of difficulties
%replaces a simpler formulation in earlier SMT-COMPs that didn't take
%into account the time solvers take.  This calculation of difficulty
%recognizes that problems requiring more time by many solvers are
%more difficult problems.  The logarithm is used to
%mark a larger change in difficulty (given a corresponding increase in
%solver average time) at smaller time scales than at higher ones (if $A=1$,
%difficulty is 0.5; at $A=1.7$, difficulty is 1.0; but a difficulty of 2.0
%requires that $A=4$); the square is used to flatten out this curve sightly
%at the low end.  
%%%Figure~\ref{difficultyplot} shows this curve.
%%%
%%%\begin{figure}
%%%  \centerline{\epsffile{difficultyplot.eps}}
%%%  \caption{The shape of the difficulty assignment curve.}
%%%  \label{difficultyplot}
%%%\end{figure}
%
%When 5 or more solvers are used for the
%calculation, the maximum and the minimum times are dropped from the calculation of the average.  Solvers giving an
%incorrect answer are not counted in the average; solvers crashing,
%timing out, or giving an unknown result are considered as taking 30
%minutes (which, with the above formula, pulls the difficulty
%toward~5.0).  If there are available solvers, but no average is
%defined under these rules, the difficulty shall be 5.0.  For new
%divisions, where there are no available solvers to compute the
%difficulty, the difficulty will be computed using whatever means are
%available to the organizers for that purpose.
%
%The following scheme will be used to choose competition benchmarks
%within each division.  Unknown-status benchmarks from SMT-LIB are
%considered ineligible for competition and are not used.
%
%The selection
%is implemented by our benchmark selection tool, source for which will
%be available at \url{www.smtcomp.org}.
%%%% TODO: Make sure this tool is available.
%
%\begin{enumerate}
%
%\item \textbf{Check benchmarks included.} %
%  All benchmarks in category \emph{check} are included.
%
%\item \textbf{Retire very easy benchmarks.} %
%  The most difficult~300 non-check non-unknown benchmarks in each
%  division are always included, together with all benchmarks on which
%  at least one 2011 solver required more than 5~seconds.  \emph{This
%    is intended to have the effect of retiring ``very easy''
%    benchmarks that were solved by every 2011 solver in less
%    than 5~seconds, \emph{unless} doing so reduces the pool of
%    benchmarks for the division to less than~300.}
%
%\item \textbf{Retire inappropriate benchmarks.} %
%  The competition organizers will remove from the eligibility pool
%  certain SMT-LIB benchmarks that are inappropriate or uninteresting
%  for competition, or cut the size of certain benchmark families to
%  avoid their over-representation.
%
%\item\label{step:pool} \textbf{Division selection pools created.} %
%  For non-\emph{check} benchmarks, selection pools are created.
%  For benchmark families with $\leq200$ eligible,
%  non-\emph{check} benchmarks, all are added to this pool;
%  otherwise, 200 such benchmarks are added to the pool with the
%  following distribution:
%%
%  \begin{itemize}
%  \item 20 with solution \textbf{sat} and difficulty on $\left[0,1\right]$
%  \item 20 with solution \textbf{sat} and difficulty on $\left(1,2\right]$
%  \item 20 with solution \textbf{sat} and difficulty on $\left(2,3\right]$
%  \item 20 with solution \textbf{sat} and difficulty on $\left(3,4\right]$
%  \item 20 with solution \textbf{sat} and difficulty on $\left(4,5\right]$
%  \item 20 with solution \textbf{unsat} and difficulty on $\left[0,1\right]$
%  \item 20 with solution \textbf{unsat} and difficulty on $\left(1,2\right]$
%  \item 20 with solution \textbf{unsat} and difficulty on $\left(2,3\right]$
%  \item 20 with solution \textbf{unsat} and difficulty on $\left(3,4\right]$
%  \item 20 with solution \textbf{unsat} and difficulty on $\left(4,5\right]$
%  \end{itemize}
%%
%  If 20 are not available in one of these subdivisions, all that are
%  available are added, and remaining slots are reallocated to the
%  others.  This process is iterated so that it is guaranteed that 200
%  benchmarks from the benchmark family are in the selection pool, in
%  equal numbers from each subdivision, so far as possible.
%  (In cases where, \eg, there are only two available slots and
%  they can be allocated to one of three subdivisions, they are allocated
%  randomly but are guaranteed to be allocated to \emph{distinct}
%  subdivisions.)
%
%\item\textbf{Category slot allocation.} %
%  Next, 200 slots are allocated for the division as follows:\footnote{The
%  number ``200'' is a guideline, and is expected to be used.  However, the
%  SMT-COMP organizers reserve the right to reduce this to an appropriate
%  value to ensure a timely end to the competition, and may do so on a
%  per-division basis.  The category allotments, etc., will remain the same
%  proportion of the total.}
%%
%  \begin{itemize}
%  \item 170 from category \emph{industrial}
%  \item 20 from category \emph{crafted}
%  \item 10 from category \emph{random}
%  \end{itemize}
%%
%  If there are fewer than 20 (respectively, 10) \emph{crafted} or
%  \emph{random} benchmarks in the division pool, more
%  \emph{industrial} slots are allocated to make 200 total for the
%  division.  If there are too few \emph{industrial} benchmarks in the
%  division pool, more \emph{crafted} slots are allocated to make 200
%  total for the division.  (In no division are there not enough of
%  both industrial and crafted benchmarks.)
%
%\item\textbf{Category subdivision slot allocation.} %
%  For each category, given that it has $n$ slots allocated to it,
%  the slot allocation is further subdivided as follows:
%%
%  \begin{itemize}
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{sat} with difficulty on $\left[0,1\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{sat} with difficulty on $\left(1,2\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{sat} with difficulty on $\left(2,3\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{sat} with difficulty on $\left(3,4\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{sat} with difficulty on $\left(4,5\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{unsat} with difficulty on $\left[0,1\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{unsat} with difficulty on $\left(1,2\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{unsat} with difficulty on $\left(2,3\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{unsat} with difficulty on $\left(3,4\right]$
%  \item $\left\lfloor n\over10\right\rfloor$ slots for solution \textbf{unsat} with difficulty on $\left(4,5\right]$
%  \end{itemize}
%%
%  Remaining slots are allocated randomly to distinct subdivisions.
%  If there aren't enough benchmarks in the pool meeting one or more of
%  the above subdivision requirements for the category, the subdivision
%  allocation is reduced to the number available in the pool that meet
%  the requirements.  To make up the full category allotment, remaining
%  slots are allocated equally to subdivisions with enough benchmarks
%  in the pool meeting their requirements.  This process ensures that
%  all slots can be filled with benchmarks from the pool.
%
%\item\textbf{Benchmark selection.} %
%  Benchmarks from the pool are assigned randomly to slots.
%\end{enumerate}
%%
%In the end, up to 200 non-\emph{check} benchmarks per division are
%included in the competition, together with all the \emph{check}
%benchmarks.  Some divisions may have fewer than 200 non-\emph{check}
%benchmarks, in which case all of them are included using this selection
%scheme.
%
%Once the StarExec service is evaluated, and if the organizers decide that
%the service has sufficient resources, the target number of benchmarks for 
%a division may be increased above 200, with all subpopulations being enlarged
%proportionately, insofar as there are sufficient benchmarks. The target number
%of benchmarks will not be raised if that requires reducing the timeout limits
%below the values used in 2011.

@@

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on
a publicized day
will be added to the other seeding values.  This helps provide transparency,
by guaranteeing that the organizers cannot manipulate the seed in
favor of or against any particular submitted solver.  Benchmarks will also be slightly
scrambled before the competition, using a simple benchmark scrambler
seeded with the same seed as the benchmark selector.  Both the
scrambler and benchmark selector will be publicly available before the
competition.  Naturally, solvers must not rely on previously
determined identifying syntactic characteristics of competition
benchmarks in testing satisfiability (violation of this is considered
cheating).

@@

\header{Benchmark sources.} %
Benchmarks for the application track will be collected by the SMT-COMP organizers.
Any benchmark available to the organizers by June 1st will be considered eligible.

\header{Benchmark availability.} 
A first release of the application track
benchmarks will be made available May 15, 2014.
More benchmarks can be collected, until June 1.
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues. 
The final release that will be used for the competition will be posted on 
June 1.

>>>>

\title{DRAFT\\Satisfiability Modulo Theories Competition (SMT-COMP) 2014: Rules and 
Procedures}

@@

\author{
David R. Cok \\
GrammaTech, Inc. \\
Ithaca, NY (USA) \\
cok@frontiernet.net \\
\and
David Deharbe \\
Federal University of Rio Grande do Norte \\
Brazil \\
David.Deharbe@loria.fr \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
tjark.weber@it.uu.se \\ 
}

@@

\noindent Comments on this document should be emailed to the smtcomp mailing
list or, if necessary, directly to the organizers (cf. email addresses in the following section and in the authors' affiliation information).

\section{Important dates and information}
\begin{itemize}
\item May 15 - deadline for new benchmarks
\item May 25 - deadline for solver entrants
\item June 1 - deadline for final versions of solvers (but no new entrants after May 25).
\item July 17 - SMT workshop; completion of competition.
\item SMT-COMP mailing list: smt-comp@cs.nyu.edu
\item signup site for mailing list: \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\item competition web site: \url{www.smtcomp.org}  or \url{smtcomp.sourceforge.net/2014}
\end{itemize}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR,SMTCOMP2012,SMTEVAL2013}).
SMT-COMP 2014 is affiliated with the SMT workshop (\url {http://smt2014.it.uu.se/}) at the co-located meetings of CAV, IJCAR, and SAT at FLoC 2014 (\url{http://vsl2014.at/}); the competition is part of the
FLoC Olympic Games (\url{http://vsl2014.at/olympics/}).

@@

SMT-COMP 2014 will have two tracks: the conventional main track and a repeat of 2012's application track.
Within a track there are multiple divisions, where each division
uses benchmarks from a specific SMT-LIB logic (or group of logics).

The rest of this document, revised from the previous 
version,\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.}
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from the previous competition rules are the following:
\begin{itemize}
\item The competition will be run with StarExec (\url{www.starexec.org}). Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [ The demonstration unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.]
\item In the main track, we will measure comparative performance within all divisions to which at least two solvers are submitted. We will use the data to determine both sequential and parallel winners.

@@

\item We will recognize winners as measured by number of benchmarks solved; we will also recognize best open source contributions and best new entrants, as described below. 
Also, importantly, the FLoC Olympic Games is sponsoring award ceremonies (with actual medals). A selection of the SMT-COMP winners (as described below) will be awarded FLoC Olympic Games medals.

@@

All results of the competition will be made public. However, 
winners will be declared only for divisions in which at least two solvers are submitted.
Although the organizers may add other solvers for comparison purposes, only solvers that
are explicitly submitted by their authors are eligible to be designated as winners.

@@

For participation in SMT-COMP, a solver must be
uploaded to Star-Exec, made publicly available, and the organizers informed of its presence {\em and the logic divisions in which it is participating}.
A submission must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

\header{System description.} %
As part of a submission,
SMT-COMP entrants are encouraged to provide a short (1--2 pages) description of
the system.  This should include a list of all authors of the system,
 their present institutional affiliations, and any appropriate acknowledgements.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool.

\header{Other systems.} %
The organizers' intent is to promote as wide a comparison among solver options as possible.
However, if the number of solver submissions is too large for the
computational resources available to the competition, the organizers reserve the right not to
accept multiple versions of solvers from the same solver team.  The organizers also reserve the right to
include other systems of interest (such as entrants in previous competitions) in the
competition.

\header{Wrapper tools.} %
A \emph{wrapper tool} is defined as any tool
which calls one or more SMT solvers not written by the author of the wrapper
tool.  The other solvers are called the \emph{wrapped} tools.  
A wrapper tool must explicitly acknowledge any tools that it wraps.
Its system description should make clear the technical innovations  by which the
wrapper tool expects to improve on the wrapped tools.

@@

%%%%%% Make main and parallel competitions together.

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via StarExec and email to the organizers by the end of May~25, 2014 (anywhere on earth).  At that time new entrants will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until the end of June~1, 2014 (anywhere on earth).  We strongly encourage
participants to use this grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using StarExec after the original deadline on May~25.

@@

@@

Solvers will be publicly evaluated in the main and application tracks, each in several logic divisions. 
All divisions with at least two solver entrants are competition divisions, i.e., a winner will be announced. The other divisions are exhibition divisions, i.e., solvers are evaluated and results posted, but no awards are announced.

@@

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the weeks leading up to SMT 2014, from about June~2 to July~17.  Results will be
announced at SMT 2014 and in an awards ceremony scheduled by the FLoC Olympic Games coordinator.
  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

@@

\header{Input and Output.}
%
Participating solvers must read a single benchmark script presented on its standard input
channel. The script is in the concrete syntax of the SMT-LIB format,
version 2.0, though with a restricted set of commands.   A
\textbf{benchmark script} is just a text file containing a sequence of SMTLIBv2 commands in which:

@@

\noindent The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated during their submission to SMT-COMP.

@@

The main track competition uses a StarExec postprocessor (named ``SMT2 results conserv") to accumulate the results.

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given benchmark, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of CPU time and the same limit on wall clock time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
% TBD- TODO - how to constrain solvers to single processes?
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.  A
timeout scores the same as if the output is ``\texttt{unknown}''.

@@

\medskip
Solvers must respond to each command sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands. There should be no \akey{get-unsat-core} or \akey{get-proof} commands in these benchmarks.

\header{Timeouts.}
A time limit is set for the whole execution of each application
benchmark (consisting of multiple queries).  We anticipate the
timeout to be around 25~minutes (as it has been in the past).\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.}

@@

In 2012 the competition ran only some logics as competitive divisions.
In 2014 we will revert to previous practice and run each logic as a competitive division, if there are two or more solvers submitted.

\subsection{Main track}

\header{Benchmark sources.} %
Benchmark formulas for each division of the main track
will be drawn from the SMT-LIB library. The deadline for new benchmarks is
May 15. The organizers will be checking and curating these until June 1;
the organizers reserve the option to exclude new benchmarks if any prove problematic for some reason.
SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.} % 
The deadline for new benchmarks is May 15, 2014.
The benchmarks will be made available as soon as possible after that, as they are checked and curated.
The goal for completing the set of benchmarks is June 1.
The set of selected benchmarks will be published when the
competition begins.

@@

Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (\ie\ benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (\ie\ benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $(2^{31}-1)$).

@@

The benchmarks are assigned a difficulty. For 2014, the difficulty measure is the
CPU time taken by the best performing solver in the 2013 SMT-EVAL. These times will be posted 
as soon after May 15 as possible.

@@

\item \textbf{Retire very easy benchmarks.} %
  Eliminate benchmarks for which all of the current solvers in SMT-EVAL-2013
  solved the benchmark in under 5 seconds,
  \emph{unless} doing so reduces the pool of
    benchmarks for the division to less than~300.

@@

Then, for each division, the benchmarks are divided into quintiles according to the difficulty measure.
Then a random selection of N competition benchmarks is made as follows:
\begin{itemize}
\item N*40\% are taken from the top quintile, if sufficient benchmarks are available
\item additional benchmarks are taken from the second quintile, to produce N*60\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the third quintile, to produce N*75\% benchmarks, if sufficient benchmarks are available
\item additional benchmarks are taken from the fourth quintile, to produce N*90\% benchmarks, if sufficient benchmarks are available
\item any remaining needed benchmarks are taken from the bottom quintile.
\end{itemize}
Within each quintile selection,
\begin{itemize}
\item if the industrial benchmarks comprise less than 85\% of the benchmarks in that quintile, 85\% will be chosen from the industrial benchmarks (if a sufficient number are available), and the others from the population of random and crafted benchmarks;
\item if the industrial benchmarks comprise at least 85\% of the benchmarks in that quintile, the benchmarks are chosen at random from the population of industrial, random and crafted benchmarks in that quintile.
\end{itemize}

Since the organizers at this point are unsure how long a set of benchmarks may take (which will depend also on the number of solvers submitted), the competition will be run in {\em heats}. For each division, some number $N$ of benchmarks will be selected as described and randomly divided into a number of heats. The heats will be run in order. If the competition ends before all heats have completed, the remaining heats will be ignored. All benchmarks will be chosen at the beginning (that is, without reference to the results of any given heat).

@@

Pseudo-random numbers will be generated using the standard C library
function \texttt{random()}, seeded (using \texttt{srandom()}) with the
sum, modulo $2^{30}$, of the numbers provided in the system
descriptions (see Section~\ref{sec:entrants} above) by all SMT-COMP
entrants other than the organizers.  Additionally, the integer part of
the opening value of the New York Stock Exchange Composite Index on
the first day the exchange is open on or after June 2
will be added to the other seeding values.  This helps provide transparency,
by guaranteeing that the organizers cannot manipulate the seed in
favor of or against any particular submitted solver.  Benchmarks will also be slightly
scrambled before the competition, using a simple benchmark scrambler
seeded with the same seed as the benchmark selector.  Both the
scrambler and benchmark selector will be publicly available before the
competition.  Naturally, solvers must not rely on previously
determined identifying syntactic characteristics of competition
benchmarks in testing satisfiability (violation of this is considered
cheating).

@@

\header{Benchmark sources.} %
Benchmarks for the application track will be collected by the SMT-COMP organizers.
Any benchmark available to the organizers by May 15 will be considered eligible.

\header{Benchmark availability.} 
A first release of the application track
benchmarks will be made available May 15, 2014.
No additional
benchmarks will be added after this date, but benchmarks can be
modified or removed to fix possible bugs or other issues. 
The final selection that will be used for the competition will be posted when the competition begins.

\header{Benchmark demographics and selection.} 
A random selection of all the available benchmarks will be used for the competition. 
As was the case in 2012, no difficulty will be assigned to the benchmarks for the application track.
Benchmarks will be slightly scrambled before the competition, using the same scrambler and 
random seed as the main track. 
If the organizers determine that there is adequate time to complete the competition, all of the benchmarks will be used. Otherwise an unbiased random subset will be used to ensure timely completion of the competition.
